---
title: "PSTAT 131 Homework 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidymodels)
library(ggthemes)
library(yardstick)

abaloneCSV <- read.csv("/Users/joshfreitas/Desktop/pstat 131/abalone.csv")
head(abaloneCSV)
```

## Question 1: Adding `age` to the dataset
```{r}
abaloneData <- mutate(abaloneCSV, age = rings + 1.5)
glimpse(abaloneData)
abaloneData %>% 
  ggplot(aes(x = age)) +
  geom_histogram(bins = 60) +
  theme_clean()
abaloneData %>% 
  ggplot(aes(x = age)) + geom_boxplot()
```

From the histogram and boxplot, we can see that the data follows a fairly normal distribution, which is 
right-skewed. The boxplot indicates that the median age is around 11, and also shows many outliers
in the data set.

## Question 2: Splitting the data
```{r}
set.seed(3162001)

abalone_split <- initial_split(abaloneData, prop = 0.8, strata = age)
abalone_testing <- testing(abalone_split)
abalone_training <- training(abalone_split)

head(abalone_testing); head(abalone_training)
```

## Question 3: Creating a recipe to predict `age` from other predictor variables
```{r}
simple_abalone_recipe <- recipe(age ~ type + longest_shell + diameter + height + whole_weight
                                + shucked_weight + viscera_weight + shell_weight, data = abalone_training)
simple_abalone_recipe

abalone_recipe <- recipe(age ~ type + longest_shell + diameter + height + whole_weight
                                + shucked_weight + viscera_weight + shell_weight, data = abalone_training) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_interact(terms = ~ starts_with('type'):shucked_weight + longest_shell:diameter + shucked_weight:shell_weight) %>% 
  step_normalize(all_predictors())
abalone_recipe
```

We don't use `rings` to predict `age` because we have already defined `age` to be an exact function of `rings`.
If we did use `rings` in our model, we would have a predictor that will perfectly predict `age`, and thus would
create a model that is not accurate or useful.

## Question 4: Creating a `lm` linear model
```{r}
lm_model <- linear_reg() %>% 
  set_engine('lm')
```

## Question 5: Setting up a workflow
```{r}
lm_wflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(abalone_recipe)
```

## Question 6: Using `fit()` to find the predicted age
```{r}
lm_fit <- fit(lm_wflow, abalone_training)
lm_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

fakeAbalone <- data.frame(type = 'F', longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1, rings = 9)

predict(lm_fit, new_data = fakeAbalone)
```

From our predict() function, we could see that the predicted age of thus hypothetical abalone
would be **21.73369**.

## Question 7: Assessing model performance with `yardstick`
```{r}
abalone_training_res <- predict(lm_fit, new_data = abalone_training %>% select(-age))
head(abalone_training_res)

abalone_training_res <- bind_cols(abalone_training_res, abalone_training %>% select(age))
head(abalone_training_res)

abalone_training_res %>% 
  ggplot(aes(x = .pred, y = age)) +
  geom_point(alpha = 0.2) +
  geom_abline(lty = 2) + 
  theme_bw() +
  coord_obs_pred()

abalone_metrics <- metric_set(rmse, rsq, mae)
abalone_metrics(abalone_training_res, truth = age, 
                estimate = .pred)
```

Our metrics have produced a `rmse`=**2.1556964**, `rsq`=**0.5477235**, and `mae`=**1.5447322**.
Generally, the RMSE and MAE are high, indicating that the residuals of the model are high
and thus that they are spread out and at a high magnitude. 
The `rsq` value of 0.5477235 indicate that only 54.77% of the variability in age of the abalone
can be predicted by our model, which is a weak relationship. Thus, based on the predictors we were provided
and our interactions, we would say that these are not effective at predicting the age of an abalone.
